One Model May Not Fit All: Subgroup Detection Using Model-Based Recursive Partitioning

### 2023-10-24 Raschtree_MH.Rnw
Changes Caro Raschtree

We can also color the DIF items in the end node profiles according to a particular split in the tree. Here, we can see that for node 1 (split in the covariate \texttt{Gender}), items 1, 6, 7, 8, and 9 show DIF in categories B or C. As a reminder, the content of these items was:

@@

These items are (relative to the anchor items for this group comparison) easier\todo{double check Mirka} to answer for participants with female or missing gender.

\todo[inline]{Caro an Mirka: ich nehme an dass hier hohe item parameter bedeuten schwierig, niedrige leicht, dann wären diese items fur Frauen leichter, du hast es unten aber glaub genau andersrum geschrieben - evtl. weil im original Raschtree paper oder irgendwo mal easiness statt difficulty abgetragen war?}

We can also color the DIF items according to node 2 (split in the covariate \texttt{Age}). Again, item 1 (Painting by Andy Warhol) shows DIF in category B. It is (relative to the anchor items for this group comparison) easier\todo{Caro glaubt: more difficult} to answer for younger respondents (age $\leq 24$) compared to older respondents.

<<fig = TRUE, width = 9, height = 7>>=
plot(Raschtree_MH_culture, color_by_node = 2)
@

For node 4 (another split in the covariate \texttt{Age}), items 1 (Painting by Andy Warhol) and 9 (Opera not by Mozart: Aida) show DIF and are (relative to the anchor items for this group comparison) easier\todo{Caro glaubt: more difficult to answer} for younger respondents (age $\leq 29$).

@@

\todo[inline]{my (Caros) feeling is that two colored trees are enough, would leave out one of the colored trees based on age, and say above: We can also color the DIF items in the end node profiles according to a particular split in the tree. We will illustrate this with two examples. First, we can see that for node 1...}

@@

The displayed information again shows that some items show only negligible DIF of category A in all nodes, while other items show medium (category B) or large (category C) DIF in some or all nodes. The values of the Mantel-Haenszel effect size measure are negative when \todo{Caro an Mirka: ich rate hier völlig aus dem nichts, bitte korrigieren} when the item difficulty is lower in the right daughter node than in the left daughter node????. For example, as we already saw from the first colored tree plot above, items 1, 6, 7, 8, and 9 have medium or large DIF (DIF categories B and C) and are  easier\todo{double check Mirka} to answer (relative to the anchor items for this group comparison) for participants with female or missing gender, which correspond to the right daughter node of node 1. Item 4 also has a negative sign, indicating that the item difficulty is lower for participants with female or missing gender. However, this item has a smaller DIF effect size. Thus it falls into DIF category A (negligible) and was not relevant for the split in \texttt{Gender}.

>>>>

We can also color the DIF items in the end node profiles according to a particular split in the tree. We will illustrate this with two examples. 
First, we can see that for node 1 (split in the covariate \texttt{Gender}), items 1, 6, 7, 8, and 9 show DIF in categories B or C. As a reminder, the content of these items was:

@@

These items are (relative to the anchor items for this group comparison) easier or less difficult to answer for participants with female or missing gender.

% We can also color the DIF items according to node 2 (split in the covariate \texttt{Age}). Again, item 1 (Painting by Andy Warhol) shows DIF in category B. It is (relative to the anchor items for this group comparison) more difficult to answer for younger respondents (age $\leq 24$) compared to older respondents.

% <<fig = TRUE, width = 9, height = 7>>=
% plot(Raschtree_MH_culture, color_by_node = 2)
% @

For node 4 (split in the covariate \texttt{Age}), items 1 (Painting by Andy Warhol) and 9 (Opera not by Mozart: Aida) show DIF and are (relative to the anchor items for this group comparison) more difficult to answer for younger respondents (age $\leq 29$), and easier or less difficult to answer for respondents $>29$.

@@

@@

The displayed information again shows that some items show only negligible DIF of category A in all nodes, while other items show medium (category B) or large (category C) DIF in some or all nodes. The values of the Mantel-Haenszel effect size measure are negative when the item difficulty is lower in the right daughter node than in the left daughter node. For example, as we already saw from the first colored tree plot above, items 1, 6, 7, 8, and 9 have medium or large DIF (DIF categories B and C) and are  easier to answer (relative to the anchor items for this group comparison) for participants with female or missing gender, which correspond to the right daughter node of node 1. % Item 4 also has a negative sign, indicating that the item difficulty is lower for participants with female or missing gender. However, this item has a smaller DIF effect size. Thus it falls into DIF category A (negligible) and was not relevant for the split in \texttt{Gender}.
